---
title: " "
author: "by Sarah Curry"
date: "19 November 2024"
output:
  html_document:
    toc: TRUE
    theme: united
    number_sections: FALSE
    toc_float: TRUE
    toc_depth: 2
    code_folding: hide
bibliography: references.bib
---

<style type="text/css">
  body{
  font-size: 12pt;
  font-family: "Century Old Style", roman;
}

  h1, h2, h3{
  font-family: "Elephant", italic;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(tigris_use_cache = TRUE,
        width = 50)
setwd("C:/LocalDocuments/GitHub/MUSA6310_HW02")

library(dplyr)
library(ggiraph)
library(ggplot2)
library(kableExtra)
library(knitr)
library(mapview)
library(scales)
library(sf)
library(tidycensus)
library(tigris)

us <- states(cb = TRUE)%>%
  st_transform(crs = 5070)
water <- st_as_sfc(st_bbox(us))
us <- us %>% 
  filter(as.numeric(GEOID) <= 56) 
pacounty <- counties(state = 42) %>% 
  select(NAME) %>%
  st_transform(crs = 5070)

acs_variable_list.2022 <- load_variables(2022, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

```
# Where's the Remote (Work)? Where to Find High Salaries and Low Costs of Living

In 2022, the median earnings per full-time worker in the United States was $60,070 [@nationalmedianearning]. 
Increasingly, the constraints of the economy have prompted everyone to consider 
how they can make more money while simultaneously spending less. 

One common strategy in this era of increasing viability for remote work includes 
working at companies that are 
based in places or regions where a higher income is more likely. In order to further maximize their 
paycheck, employees like you should look to live in places where
they can rent or own cheaper housing or where there is a low cost of living. 

So where should you look for a job and where should you live? Consider first 
what kind of salary you could be making. Here are the Median
Household Incomes in the 3 largest Metropolitan Statistical Areas per time zone 
in the United States. 

```{r fig_one_data, message=FALSE, warning=FALSE}

#https://geodata.bts.gov/datasets/usdot::time-zones/explore
timezone_us <- read_sf(dsn = "data/NTAD_Time_Zones_873827312334388986.geojson") %>%
  filter(zone == "Eastern" | zone == "Central" | zone == "Mountain" | zone == "Pacific") %>%
  st_transform(crs = 5070) %>%
  select(zone)

msa_mhi_22 <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area",
                     year = 2022, 
                     variables = c(pop = "B01003_001", mhi = "B19013_001", wfh = "B99087_005"),#median household income
                     geometry = TRUE,
                     output = "wide"
) 

#match msa to time zone
msa_ctr <- msa_mhi_22 %>%
  st_transform(crs = 5070) %>%
  st_centroid()

msa_tz <- st_intersection(msa_ctr, timezone_us) %>%
  select(GEOID, zone) %>%
  st_drop_geometry()

#check on wfh pc
msa_mhi_22 <- msa_mhi_22 %>%
  mutate(pcentwfh = wfhE/popE)

#attach tz to each msa
msa_mhi_22 <- left_join(msa_mhi_22, msa_tz, by = "GEOID")

#average msa mhi per time zone
msa_mhi_22_avg <- msa_mhi_22 %>%
  group_by(zone) %>%
  summarize(mhiavg = mean(mhiE))

#grab top 3 per tz
msa_mhi_22_eastern <- msa_mhi_22 %>%
  filter(zone == "Eastern") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22_central <- msa_mhi_22 %>%
  filter(zone == "Central") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22_mountain <- msa_mhi_22 %>%
  filter(zone == "Mountain") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22_pacific <- msa_mhi_22 %>%
  filter(zone == "Pacific") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22 <- rbind(msa_mhi_22_eastern, msa_mhi_22_central)
msa_mhi_22 <- rbind(msa_mhi_22, msa_mhi_22_mountain)
msa_mhi_22 <- rbind(msa_mhi_22, msa_mhi_22_pacific)

#variables for each msa and mhi
chicagomhi <- 85087
dallasmhi <- 83398
houstonmhi <- 78061
newyorkmhi <- 93610
dcmhi <- 119803
philadelphiamhi <- 85555
phoenixmhi <- 79935
denvermhi <- 96920
saltlakecitymhi <- 90277
lamhi <- 89105
sanfranciscomhi <- 129315
riversidemhi <- 81041

```

```{r fig_one_map, message=FALSE, warning=FALSE}

ggplot() +
  geom_sf(data = timezone_us %>% filter(zone == "Eastern"), color = "transparent", fill = "#8FC1B5", alpha = 0.5) +
  geom_sf(data = timezone_us %>% filter(zone == "Central"), color = "transparent", fill = "#589A8D", alpha = 0.5) +
  geom_sf(data = timezone_us %>% filter(zone == "Mountain"), color = "transparent", fill = "#007566", alpha = 0.5) +
  geom_sf(data = timezone_us %>% filter(zone == "Pacific"), color = "transparent", fill = "#146551", alpha = 0.5) +
  geom_sf(data = us, color = "white", fill = "transparent", size = 2) +
  geom_sf(data = msa_mhi_22, color = "#FF7300", alpha  = 0.5, aes(size = MHI)) +
  annotate("text", x = 1880000, y = 2400000, label = "New York City Metro Area\n$93,610", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 1880000, y = 2000000, label = "Philadelphia Metro Area\n$85,555", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = 1400000, y = 1850000, label = "DC Metro Area\n$119,803", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 550000, y = 2050000, label = "Chicago Metro Area\n$85,087", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 30000, y = 1022000, label = "Dallas-Ft. Worth Metro Area\n$83,398", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = 160000, y = 680000, label = "Houston Metro Area\n$78,061", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = -800000, y = 1650000, label = "Denver Metro Area\n$96,920", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1500000, y = 2300000, label = "Salt Lake City Metro Area\n$90,277", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1530000, y = 1070000, label = "Phoenix Metro Area\n$79,935", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1880000, y = 1650000, label = "Riverside Metro Area\n$81,041", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1980000, y = 1300000, label = "Los Angeles Metro Area\n$89,105", color = "#332E2B", size = 2.5, hjust = 1) + 
  annotate("text", x = -2380000, y = 1900000, label = "San Francisco Metro Area\n$129,315", color = "#332E2B", size = 2.5, hjust = 1) + 
  theme(legend.position = "none",
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()) +
  xlim(-3500000, 2900000) +
  ylim(350000, 3400000) 


```

Why look at the largest MSA's? As more and more workplaces consider going back to work,
the larger MSA's are more likely to have a greater number of remote jobs available
and are also likely to be the most urban and easy to get to in the event of a work schedule
that requires going into the office from time to time. Let's assume that remote 
employees would prefer to live and work in the same timezone for ease of completing 
assignments and attending meetings.

## Living Comfortably

The cost of living in a given region can be measured in various ways. MIT examines 
the summed monthly and annual budget that includes food, childcare, transportation, 
and housing costs. Another method to measure comfortability is by calculating if the 
monthly housing costs are less than 30% of a family's household income.

> A cost-burdened household pays more than 30% of its income on rent and utilities 
and is therefore less able to spend remaining income on other necessities, 
emergencies, or at their discretion. 

See Methodology below for more information about the calculations in this article.

```{r col_data, message=FALSE, warning=FALSE, include=FALSE}

county_mgr_msmoc_22 <- get_acs(geography = "county",
                     year = 2022, 
                     variables = c(mgr = "B25064_001", msmoc = "B25094_001"), #GROSS RENT - SMOC
                     geometry = TRUE,
                     state = c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE",
                               "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", 
                               "KY", "LA", "ME", "MD", "MA", "MI", "MS", "MO",
                               "MN", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", 
                               "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 
                               "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV",
                               "WI", "WY"),
                     output = "wide"
)

#take cheaper of rent vs homeowning
county_ls_22 <- county_mgr_msmoc_22 %>%
  mutate(cheapest = pmin(mgrE, msmocE)) %>%
  mutate(buy_rent = case_when(mgrE == cheapest ~ "Rent",
                              msmocE == cheapest ~ "Buy")) %>%
  mutate(livsal = (cheapest/.30)*12)

#time zone per county
county_ctr <- county_mgr_msmoc_22 %>%
  st_transform(crs = 5070) %>%
  st_centroid()

county_tz <- st_intersection(county_ctr, timezone_us) %>%
  select(GEOID, zone) %>%
  st_drop_geometry()

county_ls_22 <- left_join(county_ls_22, county_tz, by = "GEOID")

```

## Eastern Region

If you work remotely out of New York City, DC, or Philadelphia, where in the Eastern
time zone can you live comfortably? 

```{r eastern_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_eastern <- county_ls_22 %>%
  filter(zone == "Eastern") %>%
  mutate(pc_phl_mhi = philadelphiamhi/livsal,
         pc_nyc_mhi = newyorkmhi/livsal,
         pc_dc_mhi = dcmhi/livsal,
         NAME = gsub("'s", "", NAME)) %>%
  st_transform(crs = 5070) %>% 
  mutate(tooltip = paste(NAME, dollar(livsal), sep = ": "))

county_med_livsal_eastern <- median(county_ls_22_eastern$livsal, na.rm = TRUE)

```

```{r eastern_col_map, message=FALSE, warning=FALSE}

map_eastern <- ggplot() +
  geom_sf(data = county_ls_22_eastern, aes(fill = livsal), color = "transparent") +
  geom_sf_interactive(data = county_ls_22_eastern, aes(tooltip = tooltip, data_id = NAME),  # Add interactivity for tooltips and data_id
                      size = 0.1, color = "transparent", fill = "transparent") + 
  scale_fill_steps2(low = "#2C4D00", mid = "#F5F8EF", high = "#AA4639", midpoint = county_med_livsal_eastern, breaks = seq(5000, 90000, 10000), labels = scales::label_dollar()) +
  geom_sf(data = us, color = "white", fill = "transparent", size = 2) +
  geom_sf(data = msa_mhi_22 %>% filter(zone == "Eastern"), color = "#041E37", fill = "#041E37", size = 2, shape = 25) +
  annotate("text", x = 1880000, y = 2400000, label = "New York City Metro Area\nMHI: $93,610", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 1880000, y = 2000000, label = "Philadelphia Metro Area\nMHI: $85,555", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = 1400000, y = 1850000, label = "DC Metro Area\nMHI: $119,803", color = "#332E2B", size = 2.5, hjust = 1) +
  theme(
#    legend.position = "none",
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()
        ) +
  labs(fill = "Living Salary") +
  xlim(0, 2700000) +
  ylim(350000, 3400000)

girafe(ggobj = map_eastern) %>%
  girafe_options(
    opts_hover(css = "fill:goldenrod;"),  # Customize hover effects (change fill color on hover)
    opts_zoom(max = 10)  # Enable zooming with a maximum zoom level of 10
  )

```

We'll assume that the Median Household Income in a given area is the remote work 
salary for which you should try to negotiate. Because New York City, Philadelphia, 
and DC have some of the highest paying salaries,
you're able to live comfortably almost anywhere if you can get a job there. However, 
if you want to stretch your dollar the farthest, you'll move to where
the annual salary needed to not be cost-burdened is fairly low. 

> In the Eastern region, the median living salary is $34,200.

The most affordable areas are in rural flatlands in southern Georgia, eastern 
Kentucky, and central West Virginia. Working in New York City, Philadelphia, or DC 
and living in another popular City like Atlanta, Miami, Jacksonville, or Boston, 
will not give you as much bang for your buck.

## Central Region

```{r central_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_central <- county_ls_22 %>%
  filter(zone == "Central") %>%
  mutate(pc_chi_mhi = chicagomhi/livsal,
         pc_dallas_mhi = dallasmhi/livsal,
         pc_houston_mhi = houstonmhi/livsal,
         NAME = gsub("'", "", NAME)) %>%
  st_transform(crs = 5070) %>% 
  mutate(tooltip = paste(NAME, dollar(livsal), sep = ": "))

county_med_livsal_central <- median(county_ls_22_central$livsal, na.rm = TRUE)

```

```{r central_col_map, message=FALSE, warning=FALSE}

map_central <- ggplot() +
  geom_sf(data = county_ls_22_central, aes(fill = livsal), color = "transparent") +
  geom_sf_interactive(data = county_ls_22_central, aes(tooltip = tooltip, data_id = NAME),  # Add interactivity for tooltips and data_id
                      size = 0.1, color = "transparent", fill = "transparent") + 
  scale_fill_steps2(low = "#2C4D00", mid = "#F5F8EF", high = "#AA4639", midpoint = county_med_livsal_central, breaks = seq(5000, 90000, 10000), labels = scales::label_dollar()) +
  geom_sf(data = us, color = "white", fill = "transparent", size = 2) +
  geom_sf(data = msa_mhi_22 %>% filter(zone == "Central"), color = "#041E37", fill = "#041E37", size = 2, shape = 25) +
  annotate("text", x = 550000, y = 2050000, label = "Chicago Metro Area\nMHI: $85,087", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 30000, y = 1022000, label = "Dallas-Ft. Worth Metro Area\nMHI: $83,398", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = 160000, y = 680000, label = "Houston Metro Area\nMHI: $78,061", color = "#332E2B", size = 2.5, hjust = 0) +
  theme(
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()
        ) +
  labs(fill = "Living Salary") +
  xlim(-1000000, 1000000) +
  ylim(350000, 3400000)

girafe(ggobj = map_central) %>%
  girafe_options(
    opts_hover(css = "fill:goldenrod;"),  # Customize hover effects (change fill color on hover)
    opts_zoom(max = 10)  # Enable zooming with a maximum zoom level of 10
  )

```

Among states in the Central region, expected salaries in the major metropolitans 
are somewhat consistent. However, in contrast to the Eastern region where 
affordable counties were more clustered, in this case, counties with 
an especially low cost of living are dispersed more evenly throughout almost every 
state. This would imply that there is a "high paying" and "low living" county within 
each state, which makes some remote work restrictions and commutes more attainable.

> In the Central region, the median living salary is $30,860.

Altogether, the Central region's Living Salaries demonstrate an overall cheaper 
cost of living compared to the Eastern region. These lower expected salaries align.

## Mountain Region

```{r mountain_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_mountain <- county_ls_22 %>%
  filter(zone == "Mountain") %>%
  mutate(pc_phoenix_mhi = phoenixmhi/livsal,
         pc_denver_mhi = denvermhi/livsal,
         pc_slc_mhi = saltlakecitymhi/livsal,
         NAME = gsub("'s", "", NAME)) %>%
  st_transform(crs = 5070) %>% 
  mutate(tooltip = paste(NAME, dollar(livsal), sep = ": "))

county_med_livsal_mountain <- median(county_ls_22_mountain$livsal, na.rm = TRUE)

```

```{r mountain_col_map, message=FALSE, warning=FALSE}

map_mountain <- ggplot() +
  geom_sf(data = county_ls_22_mountain, aes(fill = livsal), color = "transparent") +
  geom_sf_interactive(data = county_ls_22_mountain, aes(tooltip = tooltip, data_id = NAME),  # Add interactivity for tooltips and data_id
                      size = 0.1, color = "transparent", fill = "transparent") + 
  scale_fill_steps2(low = "#2C4D00", mid = "#F5F8EF", high = "#AA4639", midpoint = county_med_livsal_mountain, breaks = seq(5000, 90000, 10000), labels = scales::label_dollar()) +
  geom_sf(data = us, color = "white", fill = "transparent", size = 2) +
  geom_sf(data = msa_mhi_22 %>% filter(zone == "Mountain"), color = "#041E37", fill = "#041E37", size = 2, shape = 25) +
  annotate("text", x = -800000, y = 1650000, label = "Denver Metro Area\n$96,920", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1500000, y = 2300000, label = "Salt Lake City Metro Area\n$90,277", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1530000, y = 1070000, label = "Phoenix Metro Area\n$79,935", color = "#332E2B", size = 2.5, hjust = 0) + 
  theme(
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()
        ) +
  labs(fill = "Living Salary") +
  xlim(-1800000, 0) +
  ylim(350000, 3400000)

girafe(ggobj = map_mountain) %>%
  girafe_options(
    opts_hover(css = "fill:goldenrod;"),  # Customize hover effects (change fill color on hover)
    opts_zoom(max = 10)  # Enable zooming with a maximum zoom level of 10
  )

```

States in the Mountain and Pacific regions are made up of much larger counties. The 
density of these areas should be considered when evaluating the cost of living, but 
at a glance, unlike the other regions, Mountain region states appear to be much 
less internally diverse: for instance, Montana appears to generally be more affordable than 
Arizona.

> In the Mountain region, the median living salary is $33,040.

Expected salaries in the major metropolitans vary here as widely as they do in the Eastern 
region. The Phoenix Metro Area's low expected salary and high cost of living statewide 
make seeking work there not very favorable.

## Pacific Region

```{r pacific_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_pacific <- county_ls_22 %>%
  filter(zone == "Pacific") %>%
  mutate(pc_la_mhi = lamhi/livsal,
         pc_sf_mhi = sanfranciscomhi/livsal,
         pc_riverside_mhi = riversidemhi/livsal,
         NAME = gsub("'s", "", NAME)) %>%
  st_transform(crs = 5070) %>% 
  mutate(tooltip = paste(NAME, dollar(livsal), sep = ": "))

county_med_livsal_pacific <- median(county_ls_22_pacific$livsal, na.rm = TRUE)

```

```{r pacific_col_map, message=FALSE, warning=FALSE}

map_pacific <- ggplot() +
  geom_sf(data = county_ls_22_pacific, aes(fill = livsal), color = "transparent") +
  geom_sf_interactive(data = county_ls_22_pacific, aes(tooltip = tooltip, data_id = NAME),  # Add interactivity for tooltips and data_id
                      size = 0.1, color = "transparent", fill = "transparent") + 
  scale_fill_steps2(low = "#2C4D00", mid = "#F5F8EF", high = "#AA4639", midpoint = county_med_livsal_pacific, breaks = seq(5000, 90000, 10000), labels = scales::label_dollar()) +
  geom_sf(data = us, color = "white", fill = "transparent", size = 2) +
  geom_sf(data = msa_mhi_22 %>% filter(zone == "Pacific"), color = "#041E37", fill = "#041E37", size = 2, shape = 25) +
  annotate("text", x = -1880000, y = 1650000, label = "Riverside Metro Area\nMHI: $81,041", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1980000, y = 1300000, label = "Los Angeles Metro Area\nMHI: $89,105", color = "#332E2B", size = 2.5, hjust = 1) + 
  annotate("text", x = -2380000, y = 1900000, label = "San Francisco Metro Area\nMHI: $129,315", color = "#332E2B", size = 2.5, hjust = 1) + 
  theme(
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()
        ) +
  labs(fill = "Living Salary") +
  xlim(-3300000, -900000) +
  ylim(350000, 3400000)

girafe(ggobj = map_pacific) %>%
  girafe_options(
    opts_hover(css = "fill:goldenrod;"),  # Customize hover effects (change fill color on hover)
    opts_zoom(max = 10)  # Enable zooming with a maximum zoom level of 10
  )

```

Similar to the coastal environment on the eastern side of the country, the 
oceanside counties in California are those with the highest necessary Living 
Salaries. If you can score a job in San Francisco, Riverside, or Los Angeles, 
living and working from a more affordable county in Nevada, eastern Oregon, or 
Washington would give you the most bang for your buck.

> In the Pacific region, the median living salary is $45,480.

The cost of living in the Pacific region is higher than in all the other regions, 
but the expected salaries align by reflecting a higher wage. If you were willing to cross 
state lines, then working in a place like San Francisco, but living in Esmerelda County, 
Nevada would maximize a high salary and low cost of living.

# Verdict

So which is the best region for working remotely? 

The Mountain region is not the greatest option due to the way that states with higher 
salaries also had higher costs of living and vice versa. This condition is also true of 
the Pacific region.

The Eastern region is also not a great region for remote work because 
the counties with the lowest costs of living are so far away from the high paying 
metropolitan areas. With a job that required extremely few in-person office visits, 
work in this region could be feasible.

Based on this preliminary research, the Central Region has the greatest opportunities 
for remote work. Although the expected salaries are lower in this time zone than in others, 
your commutes would be manageable because you wouldn't need to cross state 
lines in order to live in an affordable county and 
work in a high paying county.

Happy job hunting and thanks for reading! Inspiration for this article comes 
from [@mobilitymatters] and [@bestworst]

### Methodology

For every county in each time zone, the annual living wage salary was calculated using the 
monthly costs for a rented or owned home - whichever was most affordable (a.k.a. "cheapest"). 
A living salary was then determined to be the annual income a household would need 
in order to not be cost-burdened. When calculating a living wage this way, there 
is no stratification for families of different sizes. This guide is more of an 
estimate to compare one county to another. All the data in this article is from 2022.

A second edition of this article would swap the MHI as the expected salary of a remote 
worker with an actual median salary per metropolitan area for a specific career sector.

```{r msa_mhi_data, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# phl_mhi_22 <- msa_mhi_22 %>%
#   filter(GEOID == 37980)

county_mhi_22 <- get_acs(geography = "county",
                     year = 2022, 
                     variables = c(pop = "B01003_001", mhi = "B19013_001", wfh = "B99087_005"),#median household income
                     geometry = TRUE,
                     output = "wide"
) 

county_mhi_22 <- county_mhi_22 %>%
  mutate(pcentwfh = wfhE/popE)

quantile(msa_mhi_22_cnt$MHI)

```

```{r msa_mhi_chart, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

ggplot() +
  geom_point(data = msa_mhi_22, aes(x = mhiE, y = pcentwfh), color = "darkolivegreen") +
  scale_x_continuous(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_percent()) +
  geom_vline(data = msa_mhi_22_cnt, aes(xintercept = quantile(msa_mhi_22_cnt$MHI)), color = "goldenrod",
             size = 0.5) +
  # geom_text(data = msa_mhi_22_cnt, aes(x = mhiE + 3000, y = 100, label = "Philadelphia MSA"),
  #           color = "goldenrod", angle = 90) +
  labs(title = "2022 Population vs Median Household Income",
       subtitle = "US Metropolitan Statistical Areas",
       x = "Median Household Income",
       y = " ",
#       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_minimal()

```

```{r trash_text, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# In 2022, the Median Household Income in the Metropolitan Statistical Area of 
# Philadelphia, PA was $85,555. Compared to the MHI of other MSAs in the United 
# States,

# The Philadelphia MSAs MHI is in the 93rd percentile of all MSAs. Theoretically,
# securing a remote work job that was based in Philadelphia allows for a large 
# amount of mobility. Across the country, a household earning that annual 
# income could likely rent in most Counties that orbit an MSA with a lower Cost of 
# Living without risk of being cost-burdened. 

```

```{r eval=FALSE, include=FALSE}

  case_when(pc_phl_mhi >= 2.0 ~ "Twice As Comfortable",
                               pc_phl_mhi < 2.0 & pc_phl_mhi >= 1.5 ~ "Nearly Twice as Comfortable",
                               pc_phl_mhi < 1.5 & pc_phl_mhi > 1.0 ~ "More than Comfortable",
                               pc_phl_mhi == 1.0 ~ "Equally As Comfortable",
                               pc_phl_mhi < 1.0 & pc_phl_mhi >= 0.8 ~ "Uncomfortable",
                               pc_phl_mhi < 0.8 & pc_phl_mhi >= 0.5 ~ "Moderately Uncomfortable",
                               pc_phl_mhi < 0.5 & pc_phl_mhi >= 0.3 ~ "Very Uncomfortable",
                               pc_phl_mhi < 0.3 ~ "Extremely Uncomfortable"

            

ec_lrs_22$f_phl_mhi <- factor(ec_lrs_22$f_phl_mhi, levels = c("Twice As Comfortable",
                                                              "Nearly Twice as Comfortable",
                                                              "More than Comfortable",
                                                              "Equally As Comfortable",
                                                              "Uncomfortable",
                                                              "Moderately Uncomfortable",
                                                              "Very Uncomfortable",
                                                              "Extremely Uncomfortable"))

```

```{r col_map_pa, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  scale_fill_brewer(palette = "RdYlGn", na.value = "gray", direction = -1) +
  xlim(-81, -74.5) +
  ylim(39.5, 42.5) +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Pennsylvania",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_void() +
  theme(legend.position = "right")

```

```{r col_map_region, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Northeastern Region",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  xlim(-2250000, 2074000) +
  ylim(350000, 3400000) +
  theme_void() +
  theme(legend.position = "right")
```

```{r col_map, eval=FALSE, message=TRUE, warning=TRUE, include=FALSE}

ggplot() +
  geom_sf(data = us, color = "white", fill = "gray90") +
  geom_sf(data = ec_lrs_22, aes(fill = livsal), color = "transparent") +
  theme(legend.position = "none",
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()) +
  xlim(-2250000, 2074000) +
  ylim(350000, 3400000) 

```

# References
