---
title: "The Best and Worst Places to Remote Work"
author: "Sarah Curry"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    theme: united
    number_sections: FALSE
    toc_float: TRUE
    toc_depth: 2
    code_folding: hide
#bibliography: references.bib
---

<style type="text/css">
  body{
  font-size: 12pt;
  font-family: "Century Old Style", roman;
}

  h1, h2, h3{
  font-family: "Congenial Black", italic;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(tigris_use_cache = TRUE,
        width = 50)
setwd("C:/LocalDocuments/GitHub/MUSA6310_HW02")

library(dplyr)
library(ggplot2)
library(kableExtra)
library(knitr)
library(mapview)
library(scales)
library(sf)
library(tidycensus)
library(tigris)

us <- states(cb = TRUE)%>%
  st_transform(crs = 5070)
water <- st_as_sfc(st_bbox(us))
us <- us %>% 
  filter(as.numeric(GEOID) <= 56) 
pacounty <- counties(state = 42) %>% 
  select(NAME) %>%
  st_transform(crs = 5070)

acs_variable_list.2022 <- load_variables(2022, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

```
# Where's the Remote (Work)?

In this era of increasing interest in remote work, folks who are entering the job
market or looking for a new position may aim to get hired at companies that are 
based in places where they can make a high income. In order to maximize their 
paycheck, they're looking to live in places where
they can rent or own housing where there is a low cost of living - 
so where should they look for a job and where should they live? Here are the Median
Household Incomes in the 3 largest Metropolitan Statistical Areas per time zone 
in the United States. 

```{r fig_one_data, message=FALSE, warning=FALSE}

#https://geodata.bts.gov/datasets/usdot::time-zones/explore
timezone_us <- read_sf(dsn = "data/NTAD_Time_Zones_873827312334388986.geojson") %>%
  filter(zone == "Eastern" | zone == "Central" | zone == "Mountain" | zone == "Pacific") %>%
  st_transform(crs = 5070) %>%
  select(zone)

msa_mhi_22 <- get_acs(geography = "metropolitan statistical area/micropolitan statistical area",
                     year = 2022, 
                     variables = c(pop = "B01003_001", mhi = "B19013_001", wfh = "B99087_005"),#median household income
                     geometry = TRUE,
                     output = "wide"
) 

#match msa to time zone
msa_ctr <- msa_mhi_22 %>%
  st_transform(crs = 5070) %>%
  st_centroid()

msa_tz <- st_intersection(msa_ctr, timezone_us) %>%
  select(GEOID, zone) %>%
  st_drop_geometry()

#check on wfh pc
msa_mhi_22 <- msa_mhi_22 %>%
  mutate(pcentwfh = wfhE/popE)

#attach tz to each msa
msa_mhi_22 <- left_join(msa_mhi_22, msa_tz, by = "GEOID")

#average msa mhi per time zone
msa_mhi_22_avg <- msa_mhi_22 %>%
  group_by(zone) %>%
  summarize(mhiavg = mean(mhiE))

#grab top 3 per tz
msa_mhi_22_eastern <- msa_mhi_22 %>%
  filter(zone == "Eastern") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22_central <- msa_mhi_22 %>%
  filter(zone == "Central") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22_mountain <- msa_mhi_22 %>%
  filter(zone == "Mountain") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22_pacific <- msa_mhi_22 %>%
  filter(zone == "Pacific") %>%
  arrange(desc(popE)) %>%
  select(NAME, mhiE, zone) %>%
  rename("MSA" = "NAME", "MHI" = "mhiE") %>%
  head(3) %>%
  st_centroid()

msa_mhi_22 <- rbind(msa_mhi_22_eastern, msa_mhi_22_central)
msa_mhi_22 <- rbind(msa_mhi_22, msa_mhi_22_mountain)
msa_mhi_22 <- rbind(msa_mhi_22, msa_mhi_22_pacific)

#variables for each msa and mhi
chicagomhi <- 85087
dallasmhi <- 83398
houstonmhi <- 78061
newyorkmhi <- 93610
dcmhi <- 119803
philadelphiamhi <- 85555
phoenixmhi <- 79935
denvermhi <- 96920
saltlakecitymhi <- 90277
lamhi <- 89105
sanfranciscomhi <- 129315
riversidemhi <- 81041

```

```{r fig_one_map, message=FALSE, warning=FALSE}

ggplot() +
  geom_sf(data = timezone_us %>% filter(zone == "Eastern"), color = "transparent", fill = "#8FC1B5", alpha = 0.5) +
  geom_sf(data = timezone_us %>% filter(zone == "Central"), color = "transparent", fill = "#589A8D", alpha = 0.5) +
  geom_sf(data = timezone_us %>% filter(zone == "Mountain"), color = "transparent", fill = "#007566", alpha = 0.5) +
  geom_sf(data = timezone_us %>% filter(zone == "Pacific"), color = "transparent", fill = "#146551", alpha = 0.5) +
  geom_sf(data = us, color = "white", fill = "transparent", size = 2) +
  geom_sf(data = msa_mhi_22, color = "#FF7300", alpha  = 0.5, aes(size = MHI)) +
  annotate("text", x = 1880000, y = 2400000, label = "New York City Metro Area\n$93,610", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 1880000, y = 2000000, label = "Philadelphia Metro Area\n$85,555", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = 1400000, y = 1850000, label = "DC Metro Area\n$119,803", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 550000, y = 2050000, label = "Chicago Metro Area\n$85,087", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 30000, y = 1022000, label = "Dallas-Ft. Worth Metro Area\n$83,398", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = 160000, y = 680000, label = "Houston Metro Area\n$78,061", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = -800000, y = 1650000, label = "Denver Metro Area\n$96,920", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1500000, y = 2300000, label = "Salt Lake City Metro Area\n$90,277", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1530000, y = 1070000, label = "Phoenix Metro Area\n$79,935", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1880000, y = 1650000, label = "Riverside Metro Area\n$81,041", color = "#332E2B", size = 2.5, hjust = 0) + 
  annotate("text", x = -1980000, y = 1300000, label = "Los Angeles Metro Area\n$89,105", color = "#332E2B", size = 2.5, hjust = 1) + 
  annotate("text", x = -2380000, y = 1900000, label = "San Francisco Metro Area\n$129,315", color = "#332E2B", size = 2.5, hjust = 1) + 
  theme(legend.position = "none",
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()) +
  xlim(-3500000, 2900000) +
  ylim(350000, 3400000) 


```

Why look at the largest MSA's? As more and more workplaces consider going back to work,
the larger MSA's are more likely to have a greater number of remote jobs available
and are also likely to be the most urban and easy to get to in the event of a work schedule
that requires going into the office from time to time. Let's assume that remote 
employees would prefer to live and work in the same timezone for ease of completing 
assignments and attending meetings.

## Living Comfortably

The cost of living in a given region can be measured in various ways. MIT examines 
the summed monthly and annual budget that includes food, childcare, transportation, 
and housing costs. Another method to measure comfortability is by calculating if the 
monthly housing costs are less than 30% of a family's household income.

> A cost-burdened household pays more than 30% of its income on rent and utilities. 

### Methodology

For every county in each time zone, the annual living wage salary was calculated using the 
monthly costs for a rented or owned home - whichever was most affordable (a.k.a. "cheapest").

```{r col_data, message=FALSE, warning=FALSE, include=FALSE}

county_mgr_msmoc_22 <- get_acs(geography = "county",
                     year = 2022, 
                     variables = c(mgr = "B25064_001", msmoc = "B25094_001"), #GROSS RENT - SMOC
                     geometry = TRUE,
                     state = c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE",
                               "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", 
                               "KY", "LA", "ME", "MD", "MA", "MI", "MS", "MO",
                               "MN", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", 
                               "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 
                               "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV",
                               "WI", "WY"),
                     output = "wide"
)

#take cheaper of rent vs homeowning
county_ls_22 <- county_mgr_msmoc_22 %>%
  mutate(cheapest = pmin(mgrE, msmocE)) %>%
  mutate(buy_rent = case_when(mgrE == cheapest ~ "Rent",
                              msmocE == cheapest ~ "Buy")) %>%
  mutate(livsal = (cheapest/.30)*12)

#time zone per county
county_ctr <- county_mgr_msmoc_22 %>%
  st_transform(crs = 5070) %>%
  st_centroid()

county_tz <- st_intersection(county_ctr, timezone_us) %>%
  select(GEOID, zone) %>%
  st_drop_geometry()

county_ls_22 <- left_join(county_ls_22, county_tz, by = "GEOID")

```

## Eastern Region

If you work remotely out of New York City, DC, or Philadelphia, where in the Eastern
time zone can you live comfortably? 

```{r eastern_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_eastern <- county_ls_22 %>%
  filter(zone == "Eastern") %>%
  mutate(pc_phl_mhi = philadelphiamhi/livsal,
         pc_nyc_mhi = newyorkmhi/livsal,
         pc_dc_mhi = dcmhi/livsal) %>%
  st_transform(crs = 5070)

```

```{r eastern_col_map, message=FALSE, warning=FALSE}

ggplot() +
#  geom_sf(data = timezone_us %>% filter(zone == "Eastern"), color = "transparent", fill = "#8FC1B5", alpha = 0.5) +
  geom_sf(data = county_ls_22_eastern, aes(fill = livsal), color = "transparent") +
  scale_fill_steps(low = "#AA4639", high = "#FFDBAA", breaks = seq(15000, 90000, 10000), labels = scales::label_dollar()) +
  geom_sf(data = us, color = "white", fill = "transparent", size = 2) +
  geom_sf(data = msa_mhi_22 %>% filter(zone == "Eastern"), color = "#FF7300", fill = "#FF7300", size = 2, shape = 25) +
  annotate("text", x = 1880000, y = 2400000, label = "New York City Metro Area\n$93,610", color = "#332E2B", size = 2.5, hjust = 1) +
  annotate("text", x = 1880000, y = 2000000, label = "Philadelphia Metro Area\n$85,555", color = "#332E2B", size = 2.5, hjust = 0) +
  annotate("text", x = 1400000, y = 1850000, label = "DC Metro Area\n$119,803", color = "#332E2B", size = 2.5, hjust = 1) +
  theme(
#    legend.position = "none",
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()
        ) +
  xlim(-1500000, 2900000) +
  ylim(350000, 3400000)

```


## Central Region

```{r central_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_central <- county_ls_22 %>%
  filter(zone == "Central") %>%
  mutate(pc_chi_mhi = chicagomhi/livsal,
         pc_dallas_mhi = dallasmhi/livsal,
         pc_houston_mhi = houstonmhi/livsal)

```

## Mountain Region

```{r mountain_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_mountain <- county_ls_22 %>%
  filter(zone == "Mountain") %>%
  mutate(pc_phoenix_mhi = phoenixmhi/livsal,
         pc_denver_mhi = denvermhi/livsal,
         pc_slc_mhi = saltlakecitymhi/livsal)

```

## Pacific Region

```{r pacific_col_data, message=FALSE, warning=FALSE, include=FALSE}

county_ls_22_pacific <- county_ls_22 %>%
  filter(zone == "Pacific") %>%
  mutate(pc_la_mhi = lamhi/livsal,
         pc_sf_mhi = sanfranciscomhi/livsal,
         pc_riverside_mhi = riversidemhi/livsal)

```

```{r col_map, eval=FALSE, message=TRUE, warning=TRUE, include=FALSE}

ggplot() +
  geom_sf(data = us, color = "white", fill = "gray90") +
  geom_sf(data = ec_lrs_22, aes(fill = livsal), color = "transparent") +
  theme(legend.position = "none",
        panel.background = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank()) +
  xlim(-2250000, 2074000) +
  ylim(350000, 3400000) 

```


Twice as Comfortable refers to a County where the Cost of Living is less than 
half of the Median Household Income in Philadelphia. Extremely Uncomfortable means
that the household living that County, with the salary of the Philadelphia Median 
Household Income, would be in poverty.

```{r col_map_region, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Northeastern Region",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  xlim(-2250000, 2074000) +
  ylim(350000, 3400000) +
  theme_void() +
  theme(legend.position = "right")
```

In the Northeastern region, the only area where a household would not be comfortable 
is in the DC/Northern Virginia area. Throughout most of Pennsylvania, a household
would be more than comfortable given the cost of living and price of rents per County.

```{r col_map_pa, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
ggplot() +
  geom_sf(data = water, fill = "steelblue") +
  geom_sf(data = ec_lrs_22, aes(fill = f_phl_mhi), color = "transparent") +
  scale_fill_brewer(palette = "RdYlGn", na.value = "gray", direction = -1) +
  xlim(-81, -74.5) +
  ylim(39.5, 42.5) +
  geom_sf(data = us, fill = "transparent", color = "steelblue", size = 8) +
  labs(title = "2022 Remote Work Bargains for Renters",
       subtitle = "Pennsylvania",
       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_void() +
  theme(legend.position = "right")

```

Inspiration for this article comes from [@mobilitymatters] and [@bestworst]

### References

# Bike Rack
In 2022, the Median Household Income in the Metropolitan Statistical Area of 
Philadelphia, PA was $85,555. Compared to the MHI of other MSAs in the United 
States,

```{r msa_mhi_data, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# phl_mhi_22 <- msa_mhi_22 %>%
#   filter(GEOID == 37980)

county_mhi_22 <- get_acs(geography = "county",
                     year = 2022, 
                     variables = c(pop = "B01003_001", mhi = "B19013_001", wfh = "B99087_005"),#median household income
                     geometry = TRUE,
                     output = "wide"
) 

county_mhi_22 <- county_mhi_22 %>%
  mutate(pcentwfh = wfhE/popE)

quantile(msa_mhi_22_cnt$MHI)

```


```{r msa_mhi_chart, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

ggplot() +
  geom_point(data = msa_mhi_22, aes(x = mhiE, y = pcentwfh), color = "darkolivegreen") +
  scale_x_continuous(labels = scales::label_dollar()) +
  scale_y_continuous(labels = scales::label_percent()) +
  geom_vline(data = msa_mhi_22_cnt, aes(xintercept = quantile(msa_mhi_22_cnt$MHI)), color = "goldenrod",
             size = 0.5) +
  # geom_text(data = msa_mhi_22_cnt, aes(x = mhiE + 3000, y = 100, label = "Philadelphia MSA"),
  #           color = "goldenrod", angle = 90) +
  labs(title = "2022 Population vs Median Household Income",
       subtitle = "US Metropolitan Statistical Areas",
       x = "Median Household Income",
       y = " ",
#       fill = "Percentage of Philadelphia MHI",
       caption = "Source: 2022 ACS, U.S. Census Bureau") +
  theme_minimal()

```

The Philadelphia MSA's MHI is in the 93rd percentile of all MSAs. Theoretically,
securing a remote work job that was based in Philadelphia allows for a large 
amount of mobility. Across the country, a household earning that annual 
income could likely rent in most Counties that orbit an MSA with a lower Cost of 
Living without risk of being cost-burdened. 

```{r eval=FALSE, include=FALSE}

  case_when(pc_phl_mhi >= 2.0 ~ "Twice As Comfortable",
                               pc_phl_mhi < 2.0 & pc_phl_mhi >= 1.5 ~ "Nearly Twice as Comfortable",
                               pc_phl_mhi < 1.5 & pc_phl_mhi > 1.0 ~ "More than Comfortable",
                               pc_phl_mhi == 1.0 ~ "Equally As Comfortable",
                               pc_phl_mhi < 1.0 & pc_phl_mhi >= 0.8 ~ "Uncomfortable",
                               pc_phl_mhi < 0.8 & pc_phl_mhi >= 0.5 ~ "Moderately Uncomfortable",
                               pc_phl_mhi < 0.5 & pc_phl_mhi >= 0.3 ~ "Very Uncomfortable",
                               pc_phl_mhi < 0.3 ~ "Extremely Uncomfortable"

            

ec_lrs_22$f_phl_mhi <- factor(ec_lrs_22$f_phl_mhi, levels = c("Twice As Comfortable",
                                                              "Nearly Twice as Comfortable",
                                                              "More than Comfortable",
                                                              "Equally As Comfortable",
                                                              "Uncomfortable",
                                                              "Moderately Uncomfortable",
                                                              "Very Uncomfortable",
                                                              "Extremely Uncomfortable"))

```

